---
title: "Formative 1"
author: "Len Metson"
date: "12th October 2023"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = FALSE)
# we set echo=FALSE to suppress code such that it by default does not appear throughout the document. 
```

```{r}
library(magrittr)
```


# Exercise 1

My forked repository can be found at the URL: [https://github.com/lenmetson/my472_formative_1](https://github.com/lenmetson/my472_formative_1). 

# Exercise 2

```{r}
data("Orange") 

tibble::as_tibble(Orange)
```

The dataset `Orange` is tidy with tree-age our unit of analysis. Each observation (`Tree` at `age`) is represented in it's own row, and each variable (`Tree`, `age`, `circumference`) has it's own column. We kan see this when we print the tibble using the `head()` function: 

```{r}
head(Orange)
```

I now create a wide version of this dataset to make it untidy. Specifically, we will represent `Tree` in the rows of the table and `age` in the columns. Values are taken from the `circumfence` column. In order to do this, I use the `tidyr::pivot_wider()` function, specifying the `names_from` parameter as `age` so that this appears in the column names, and the `values_from` parameter as `circumference`. 

```{r}
Orange <- Orange %>%
  tidyr::pivot_wider(
    names_from = age,
    values_from = circumference
  )
```

We can see by running the `head()` function again that the data is now untidy, with each column representing an age, and each row representing a tree. 

```{r}
head(Orange)
```

# Exercise 3

```{r}
data("warpbreaks")
```

```{r, echo = TRUE}
warpbreaks_breaks_by_tension <- warpbreaks %>% # Name new object 
  dplyr::group_by(tension) %>% # Group the data by tension
  dplyr::summarise(average_breaks = mean(breaks, na.rm = TRUE)) # For each new group, take the mean number of breaks

# Add transformation of a numeric variable after summarise()
  
```

```{r}
head(warpbreaks_breaks_by_tension)
```

By transforming the data in this way, we can more easily compare the differences between different tensions. We can see that there are more breaks on average when tension is low, than when tension is high for example. 

# Code Appendix

```{r ref.label=knitr::all_labels(), echo=TRUE} 

```
