---
title: "Formative 1"
author: "Len Metson"
date: "12th October 2023"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = FALSE)
# we set echo=FALSE to suppress code such that it by default does not appear throughout the document. 
```

```{r}
library(magrittr)
```


# Exercise 1

My forked repository can be found at the URL: [https://github.com/lenmetson/my472_formative_1](https://github.com/lenmetson/my472_formative_1). 

# Exercise 2

```{r}
data("Orange") 

tibble::as_tibble(Orange)
```

The dataset `Orange` is tidy with tree-age our unit of analysis. Each observation (`Tree` at `age`) is represented in it's own row, and each variable (`Tree`, `age`, `circumference`) has it's own column. We kan see this when we print the tibble using the `head()` function: 

```{r}
head(Orange)
```

I now create a wide version of this dataset to make it untidy. Specifically, we will represent `Tree` in the rows of the table and `age` in the columns. Values are taken from the `circumfence` column. In order to do this, I use the `tidyr::pivot_wider()` function, specifying the `names_from` parameter as `age` so that this appears in the column names, and the `values_from` parameter as `circumference`. 

```{r}
Orange <- Orange %>%
  tidyr::pivot_wider(
    names_from = age,
    values_from = circumference
  )
```

We can see by running the `head()` function again that the data is now untidy, with each column representing an age, and each row representing a tree. 

```{r}
head(Orange)
```

# Exercise 3

# Code Appendix

```{r ref.label=knitr::all_labels(), echo=TRUE} 

```
